Modeling and Validation of Metadata on \\ Highly-Complex Person-Level Data Sets in RDF

### 2015-01-12 paper/RDF validation requirements/RDF validation requirements.tex
paper ideas
disco constraints
incorporate feedback

\tb{why is the references section not shown? / references are not resolved}

@@

\ms{Existential Quantification on Properties} conforms to the requirement \\
\ms{R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES}.
In DL terminology the \ms{existential quantification} is alco called \ms{existential restriction}.
An existential class expression consists of a property expression and a class expression or a data range, and it contains all those individuals that are connected by the property expression to an individual that is an instance of the class expression or to literals that are in the data range.

@@

\begin{DL}
$\forall$ parentOf . Female
\end{DL}

@@

\ms{Minimum Qualified Cardinality Restrictions on Properties} corresponds to the requirements
\ms{R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES} and \ms{R-211-CARDINALITY-CONSTRAINTS}
A minimum cardinality restrictions contains all those individuals that are connected by a property to at least n different individuals/literals 
that are instances of a particular class or data range. If the class is missing, it is taken to be owl:Thing. 
If the data range is missing, it is taken to be rdfs:Literal.
For qualified cardinality restrictions, classes respective data ranges are stated.
$\leq n R. C$ is a minimum qualified cardinality restriction where $n \in \mathbb{N}$.

@@

\begin{ex}
# DSCL (DQTP):
SELECT ?s WHERE { 
    ?s %%P1%% ?v1 .
    ?s %%P2%% ?v2 .
    FILTER ( ?v1 %%OP%% ?v2 ) }
\end{ex}

@@

\ms{Literal Ranges} corresponds to the requirement
\ms{R-45-RANGES-OF-RDF-LITERAL-} \ms{VALUES}.

@@

\ms{Recommended Properties} corresponds to the requirement \\
\ms{R-72-RECOMMENDED-PROPERTIES}.

>>>>

@@

\ms{Existential Quantification on Properties} conforms to the requirement \\
\ms{R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES}.
In DL terminology the \ms{existential quantification} is also called \ms{existential restriction}.
An existential class expression consists of a property expression and a class expression or a data range, and it contains all those individuals that are connected by the property expression to an individual that is an instance of the class expression or to literals that are in the data range.

@@

\begin{DL}
$\forall$ parentOf.Female
\end{DL}

@@

\ms{Minimum Qualified Cardinality Restrictions on Properties} corresponds to the requirements
\ms{R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES} and \ms{R-211-CARDINALITY-CONSTRAINTS}
A minimum cardinality restrictions contains all those individuals that are connected by a property to at least n different individuals/literals 
that are instances of a particular class or data range. If the class is missing, it is taken to be owl:Thing. 
If the data range is missing, it is taken to be rdfs:Literal.
For qualified cardinality restrictions, classes respective data ranges are stated.
$\geq n R. C$ is a minimum qualified cardinality restriction where $n \in \mathbb{N}$.

@@

\begin{ex}
# DQTP:
SELECT ?s WHERE { 
    ?s %%P1%% ?v1 .
    ?s %%P2%% ?v2 .
    FILTER ( ?v1 %%OP%% ?v2 ) }
\end{ex}

@@

\ms{Literal Ranges} corresponds to the requirement
\ms{R-45-RANGES-OF-RDF-LITERAL-} \ms{VALUES}.
P1 is a data property (of an instance of class C1) and its literal value must be between the range of [$V_{min}$,$V_{max}$].

@@

\ms{Recommended Properties} corresponds to the requirement \\
\ms{R-72-RECOMMENDED-PROPERTIES}.
Which properties are not required but recommended within a particular context.

### 2015-01-12 paper/paper6/paper6.tex
incorporate feedback

@@

A minimum cardinality restrictions contains all those individuals that are connected by a property to at least n different individuals/literals 
that are instances of a particular class or data range. If the class is missing, it is taken to be owl:Thing. 
If the data range is missing, it is taken to be rdfs:Literal.
$\geq n R. C$ is a minimum qualified cardinality restriction where $n \in \mathbb{N}$\footnote{{\em R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES} and {\em R-211-CARDINALITY-CONSTRAINTS}}.

\textbf{{\em DISCO-C-XX}}.
A {\em disco:Questionnaire} has at least 1 {\em disco:question} relationship to {\em disco:Question}\footnote{When mapped to DL, we do not state namespace prefixes for simplicity reasons}.

@@

An exact cardinality restriction contains all those individuals that are connected by a property to exactly n different individuals that are instances of a particular class or data range. 
If the class is missing, it is taken to be owl:Thing. 
If the data range is not present, it is taken to be rdfs:Literal.
$\geq n R. C \sqcap \leq n R. C $ is an exact qualified cardinality restriction where $n \in \mathbb{N}$\footnote{{\em R-74-EXACT-QUALIFIED-CARDINALITY-ON-PROPERTIES} and {\em R-211-CARDINALITY-CONSTRAINTS}.}.

\textbf{{\em DISCO-C-XX}}. 
A {\em disco:Question} has exactly 1 {\em disco:universe} relationship to {\em disco:Universe}.

@@

A universal class expression ({\em value restriction} in DL) contains all those individuals that are connected by an object property only to individuals that are instances of a particular class\footnote{{\em R-91-UNIVERSAL-QUANTIFICATION-ON-PROPERTIES}}.

\textbf{{\em DISCO-C-XX}}.
Only {\em disco:LogicalDataSet}s can have {\em disco:aggregation} relationships to {\em qb:DataSet}s.

@@

An existential class expression ({\em existential restriction} in DL terminology) contains all those individuals that are connected by the property P to an individual x that is an instance of the class C or to literals that are in the data range DR\footnote{{\em R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES}}.

\textbf{{\em DISCO-C-XX}}. 
There must be at least 1 {\em disco:universe} relationship from {\em disco:Studies} or {\em disco:StudyGroups} to {\em disco:Universe}.

@@

A union class expression contains all individuals that are instances of at least one class $C_{i}$ for 1 $\leq$ i $\leq$ n. 
A union data range contains all tuples of literals that are contained in the at least one data range $DR_{i}$ for 1 $\leq$ i $\leq$ n.
Synonyms of {\em disjunction} are {\em union} and {\em inclusive or}\footnote{{\em R-17-DISJUNCTION-OF-CLASS-EXPRESSIONS} and {\em R-18-DISJUNCTION-OF-DATA-RANGES}}.

\textbf{{\em DISCO-C-XX}}. 
Only {\em disco:Variable}s or {\em disco:Question}s or {\em disco:RepresentedVariable}s can have {\em disco:concept} relationships to {\em skos:Concept}s.

@@

{\em Property Domains} ({\em domain restrictions on roles} in DL) restricts the domain of object and data properties.
The purpose is to declare that a given property is associated with a class. 
In OO terms this is the declaration of a member, field, attribute or association. 
$\exists R. \top \sqsubseteq C$ is the object property restriction where $R$ is the object property (role) whose domain is restricted to concept $C$\footnote{{\em R-25-OBJECT-PROPERTY-DOMAIN} and {\em R-26-DATA-PROPERTY-DOMAIN}}.
{\em Property Domain} constraints are defined for each Disco object and data property.

\textbf{{\em DISCO-C-XX}}. 
Only {\em disco:Question}s can have {\em disco:responseDomain} relationships.

@@

{\em Property Range} ({\em range restrictions on roles} in DL) restricts the range of object and data properties.
$\top \sqsubseteq \forall R . C$ is the range restriction to the object property $R$ (restricted by the concept $C$)\footnote{{\em R-28-OBJECT-PROPERTY-RANGE} and {\em R-35-DATA-PROPERTY-RANGE}}.

\textbf{{\em DISCO-C-XX}}. 
{\em disco:caseQuantity} relationships can only point to literals of the datatype {\em xsd:nonNegativeInteger}.

@@

{\em Class-Specific Property Range} restricts the range of object and data properties for individuals within a specific context (e.g. class, shape, application profile).
The values of each member property of a class may be limited by their value type, such as xsd:string or foaf:Person\footnote{{\em R-29-CLASS-SPECIFIC-RANGE-OF-RDF-OBJECTS} and {\em R-36-CLASS-SPECIFIC-RANGE-OF-RDF-LITERALS}}.

\textbf{{\em DISCO-C-XX}}. 
Only {\em disco:Question}s can have {\em disco:questionText} relationships to literals of the datatype {\em rdf:langString}.

@@

A subclass axiom ({\em concept inclusion} in DL) states that the class C1 is a subclass of the class C2 - C1 is more specific than C2\footnote{{\em R-100-SUBSUMPTION}}.

\textbf{{\em DISCO-C-XX}}. 
All {\em disco:Universe}s must also be {\em skos:Concept}s.

@@

{\em Class Equivalence} asserts that two concepts have the same instances.
While synonyms are an obvious example of equivalent concepts, in practice one more
often uses concept equivalence to give a name to complex expressions \cite{Kroetzsch2012}.
Concept equivalence is indeed subsumption from left and right ($A \sqsubseteq B$ and $B \sqsubseteq A$ implies $A \equiv B$)\footnote{{\em R-3-EQUIVALENT-CLASSES}}.

\textbf{{\em DISCO-C-XX}}. 
All {\em sio:SIO\_000367}s must also be {\em disco:Variable}s.

@@

{\em Sub Properties} state that the property P1 is a sub property of the property P2 - that is, if an individual x is connected by P1 to an individual or a literal y, then x is also connected by P2 to y.

\textbf{{\em DISCO-C-XX}}. 
If an individual x is connected by {\em disco:fundedBy} to an individual y, then x is also connected by {\em dcterms:contributor} to y.

@@

{\em Disjoint Classes} state that all of the classes are pairwise disjoint; 
that is, no individual can be at the same time an instance of these disjoint classes\footnote{{\em R-7-DISJOINT-CLASSES}}.
\textbf{{\em DISCO-C-01}}. All Disco classes are defined to be pairwise disjoint.
The following DL statements holds for each pair of Disco classes:

@@

A property is asymmetric if it is disjoint from its own inverse \cite{Kroetzsch2012}.
An object property asymmetry axiom states that the object property OP is asymmetric - that is, if an individual x is connected by OP to an individual y, then y cannot be connected by OP to x\footnote{{\em R-62-ASYMMETRIC-OBJECT-PROPERTIES}}.

\textbf{{\em DISCO-C-01}}. 
A {\em disco:Variable} may be based on a {\em disco:RepresentedVariable}.
A {\em disco:RepresentedVariable}, however, cannot be based on a {\em disco:Variable}.
This is a kind of mistake may occur as a semantically equivalent object property for the other direction may also be possible ({\em disco:basisOf}).

@@

A property is irreflexive if it is never locally reflexive \cite{Kroetzsch2012}.
An object property irreflexivity axiom states that the object property OP is irreflexive - that is, no individual is connected by OP to itself.

\textbf{{\em DISCO-C-XX}}.
Within the Disco context, {\em skos:Concept}s cannot be related via the object property {\em skos:boader} to themselves.

\begin{DL}
Concept $\sqsubseteq$ $\neg$$\exists$ broader.Self. 
\end{DL}

\textbf{{\em DISCO-C-XX}}.
Within the Disco context, {\em skos:Concept}s cannot be related via the object property {\em skos:narrower} to themselves.

@@

Exclusive or is a logical operation that outputs true whenever both inputs differ (one is true, the other is false).
Only one of multiple property groups leads to valid data\footnote{{\em R-13-DISJOINT-GROUP-OF-PROPERTIES-CLASS-SPECIFIC}}.

\textbf{{\em DISCO-C-XX}}.
Within the context of Disco, skos:Concepts can have either skos:definition (when interpreted as DDI concepts) or skos:notation and skos:prefLabel properties (when interpreted as DDI codes and categories), but not both.

@@

\section{Allowed Values}

@@

>>>>

\textcolor{blue}{multiple validation levels}

@@

A minimum cardinality restrictions contains all those individuals that are connected by a property to at least n different individuals/literals 
that are instances of a particular class or data range. If the class is missing, it is taken to be owl:Thing. 
If the data range is missing, it is taken to be rdfs:Literal.
$\geq n R. C$ is a minimum qualified cardinality restriction where $n \in \mathbb{N}$\footnote{{\em R-75-MINIMUM-QUALIFIED-CARDINALITY-ON-PROPERTIES} and {\em R-211-CARDINALITY-CONSTRAINTS}}.
\textbf{{\em DISCO-C-XX:}}
A {\em disco:Questionnaire} has at least 1 {\em disco:question} relationship to {\em disco:Question}\footnote{When mapped to DL, we do not state namespace prefixes for simplicity reasons}.

@@

An exact cardinality restriction contains all those individuals that are connected by a property to exactly n different individuals that are instances of a particular class or data range. 
If the class is missing, it is taken to be owl:Thing. 
If the data range is not present, it is taken to be rdfs:Literal.
$\geq n R. C \sqcap \leq n R. C $ is an exact qualified cardinality restriction where $n \in \mathbb{N}$\footnote{{\em R-74-EXACT-QUALIFIED-CARDINALITY-ON-PROPERTIES} and {\em R-211-CARDINALITY-CONSTRAINTS}.}.
\textbf{{\em DISCO-C-XX:}}
A {\em disco:Question} has exactly 1 {\em disco:universe} relationship to {\em disco:Universe}.

@@

A universal class expression ({\em value restriction} in DL) contains all those individuals that are connected by an object property only to individuals that are instances of a particular class\footnote{{\em R-91-UNIVERSAL-QUANTIFICATION-ON-PROPERTIES}}.
\textbf{{\em DISCO-C-XX:}}
Only {\em disco:LogicalDataSet}s can have {\em disco:aggregation} relationships to {\em qb:DataSet}s.

@@

An existential class expression ({\em existential restriction} in DL terminology) contains all those individuals that are connected by the property P to an individual x that is an instance of the class C or to literals that are in the data range DR\footnote{{\em R-86-EXISTENTIAL-QUANTIFICATION-ON-PROPERTIES}}.
\textbf{{\em DISCO-C-XX:}} 
There must be at least 1 {\em disco:universe} relationship from {\em disco:Studies} or {\em disco:StudyGroups} to {\em disco:Universe}.

@@

A union class expression contains all individuals that are instances of at least one class $C_{i}$ for 1 $\leq$ i $\leq$ n. 
A union data range contains all tuples of literals that are contained in the at least one data range $DR_{i}$ for 1 $\leq$ i $\leq$ n.
Synonyms of {\em disjunction} are {\em union} and {\em inclusive or}\footnote{{\em R-17-DISJUNCTION-OF-CLASS-EXPRESSIONS} and {\em R-18-DISJUNCTION-OF-DATA-RANGES}}.
\textbf{{\em DISCO-C-XX:}} 
Only {\em disco:Variable}s or {\em disco:Question}s or {\em disco:RepresentedVariable}s can have {\em disco:concept} relationships to {\em skos:Concept}s.

@@

{\em Property Domains} ({\em domain restrictions on roles} in DL) restricts the domain of object and data properties.
The purpose is to declare that a given property is associated with a class. 
In OO terms this is the declaration of a member, field, attribute or association. 
$\exists R. \top \sqsubseteq C$ is the object property restriction where $R$ is the object property (role) whose domain is restricted to concept $C$\footnote{{\em R-25-OBJECT-PROPERTY-DOMAIN} and {\em R-26-DATA-PROPERTY-DOMAIN}}.
{\em Property Domain} constraints are defined for each Disco object and data property.
\textbf{{\em DISCO-C-XX:}} 
Only {\em disco:Question}s can have {\em disco:responseDomain} relationships.

@@

{\em Property Range} ({\em range restrictions on roles} in DL) restricts the range of object and data properties.
$\top \sqsubseteq \forall R . C$ is the range restriction to the object property $R$ (restricted by the concept $C$)\footnote{{\em R-28-OBJECT-PROPERTY-RANGE} and {\em R-35-DATA-PROPERTY-RANGE}}. 
\textbf{{\em DISCO-C-XX:}} 
{\em disco:caseQuantity} relationships can only point to literals of the datatype {\em xsd:nonNegativeInteger}.

@@

{\em Class-Specific Property Range} restricts the range of object and data properties for individuals within a specific context (e.g. class, shape, application profile).
The values of each member property of a class may be limited by their value type, such as xsd:string or foaf:Person\footnote{{\em R-29-CLASS-SPECIFIC-RANGE-OF-RDF-OBJECTS} and {\em R-36-CLASS-SPECIFIC-RANGE-OF-RDF-LITERALS}}. 
\textbf{{\em DISCO-C-XX:}} 
Only {\em disco:Question}s can have {\em disco:questionText} relationships to literals of the datatype {\em rdf:langString}.

@@

A subclass axiom ({\em concept inclusion} in DL) states that the class C1 is a subclass of the class C2 - C1 is more specific than C2\footnote{{\em R-100-SUBSUMPTION}}.
\textbf{{\em DISCO-C-XX:}} 
All {\em disco:Universe}s must also be {\em skos:Concept}s.

@@

{\em Class Equivalence} asserts that two concepts have the same instances.
While synonyms are an obvious example of equivalent concepts, in practice one more
often uses concept equivalence to give a name to complex expressions \cite{Kroetzsch2012}.
Concept equivalence is indeed subsumption from left and right ($A \sqsubseteq B$ and $B \sqsubseteq A$ implies $A \equiv B$)\footnote{{\em R-3-EQUIVALENT-CLASSES}}.
\textbf{{\em DISCO-C-XX:}}
All {\em sio:SIO\_000367}s must also be {\em disco:Variable}s.

@@

{\em Sub Properties} state that the property P1 is a sub property of the property P2 - that is, if an individual x is connected by P1 to an individual or a literal y, then x is also connected by P2 to y. 
\textbf{{\em DISCO-C-XX:}}
If an individual x is connected by {\em disco:fundedBy} to an individual y, then x is also connected by {\em dcterms:contributor} to y.

@@

{\em Disjoint Classes} state that all of the classes are pairwise disjoint; 
that is, no individual can be at the same time an instance of these disjoint classes\footnote{{\em R-7-DISJOINT-CLASSES}}.
\textbf{{\em DISCO-C-XX:}} 
All Disco classes are defined to be pairwise disjoint.
The following DL statements holds for each pair of Disco classes:

@@

A property is asymmetric if it is disjoint from its own inverse \cite{Kroetzsch2012}.
An object property asymmetry axiom states that the object property OP is asymmetric - that is, if an individual x is connected by OP to an individual y, then y cannot be connected by OP to x\footnote{{\em R-62-ASYMMETRIC-OBJECT-PROPERTIES}}. 
\textbf{{\em DISCO-C-XX:}} 
A {\em disco:Variable} may be based on a {\em disco:RepresentedVariable}.
A {\em disco:RepresentedVariable}, however, cannot be based on a {\em disco:Variable}.
This is a kind of mistake may occur as a semantically equivalent object property for the other direction may also be possible ({\em disco:basisOf}).

@@

A property is irreflexive if it is never locally reflexive \cite{Kroetzsch2012}.
An object property irreflexivity axiom states that the object property OP is irreflexive - that is, no individual is connected by OP to itself. 
\textbf{{\em DISCO-C-XX:}}
Within the Disco context, {\em skos:Concept}s cannot be related via the object property {\em skos:boader} to themselves (DL: Concept $\sqsubseteq$ $\neg$$\exists$ broader.Self).

\textbf{{\em DISCO-C-XX:}}
Within the Disco context, {\em skos:Concept}s cannot be related via the object property {\em skos:narrower} to themselves.

@@

Exclusive or is a logical operation that outputs true whenever both inputs differ (one is true, the other is false).
Only one of multiple property groups leads to valid data\footnote{{\em R-13-DISJOINT-GROUP-OF-PROPERTIES-CLASS-SPECIFIC}}.
\textbf{{\em DISCO-C-XX:}}
Within the context of Disco, skos:Concepts can have either skos:definition (when interpreted as DDI concepts) or skos:notation and skos:prefLabel properties (when interpreted as DDI codes and categories), but not both.

@@

\subsection{Allowed Values}

@@

\subsection{Literal Ranges}

P1 is a data property (of an instance of class C1) and its literal value must be between the range of [$V_{min}$,$V_{max}$].
\textbf{{\em DISCO-C-XX:}}
{\em disco:percentage} (domain: {\em disco:CategoryStatistics}) literals must be of the datatype {\em xsd:double} whose range should be restricted to be between 0 and 100 (not expressible in DL)\footnote{{\em R-45-RANGES-OF-RDF-LITERAL-VALUES}}.

\subsection{Mathematical Operations}

Examples for {\em Mathematical Operations} are the addition of two dates, the addition of days to a start date, and statistical computations (e.g. average, mean, sum).
\textbf{{\em DISCO-C-XX:}}
The sum of {\em disco:percentage} {\em xsd:double} values of all codes (represented as {\em skos:Concept}s) of a code list ({\em skos:ConceptScheme} or {\em skos:OrderedCollection}), serving as representation of a particular {\em disco:Variable}, must exactly be 100  
(not expressible in DL)\footnote{{\em R-42-MATHEMATICAL-OPERATIONS} and {\em R-41-STATISTICAL-COMPUTATIONS}}.

\subsection{Default Values}

Default values for objects and literals are inferred automatically.
It should be possible to declare the default value for a given property, e.g. so that input forms can be pre-populated and to insert a required property that is missing in a web service call\footnote{{\em R-31-DEFAULT-VALUES-OF-RDF-OBJECTS} and {\em R-38-DEFAULT-VALUES-OF-RDF-LITERALS}}.
\textbf{{\em DISCO-C-XX:}}
The value 'true' for the property {\em disco:isPublic} ({\em xsd:boolean}) indicates that the data set ({\em disco:LogicalDataSet}) can be accessed (usually downloaded) by anyone.
Per default, access to data sets should be restricted ('false').

\subsection{Language Tag Matching}

For particular data properties, values must be stated for predefined languages (not expressible in DL)\footnote{{\em R-47-LANGUAGE-TAG-MATCHING}}.
\textbf{{\em DISCO-C-XX:}}
There must be an English variable name ({\em skos:notation}) for each {\em disco:Variable} within {\em disco:LogicalDataSet}s.

\subsection{Language Tag Cardinality}

For particular data properties, values of predefined languages must be stated for determined number of times (not expressible in DL)
\footnote{{\em R-49-RDF-LITERALS-HAVING-AT-MOST-ONE-LANGUAGE-TAG} and {\em R-48-MISSING-LANGUAGE-TAGS}}.
\textbf{{\em DISCO-C-XX:}}
There must be at least one English {\em disco:questionText} for each {\em disco:Question} within {\em disco:LogicalDataSet}s.
\textbf{{\em DISCO-C-XX:}}
There should be at most one English literal value for variable names ({\em skos:notation}, domain: {\em disco:Variable}).

\subsection{Conditional Properties}

If specific properties exist, then specific other properties also have to be present\footnote{{\em R-71-CONDITIONAL-PROPERTIES}}.
\textbf{{\em DISCO-C-XX:}}
If a {\em skos:Concept represents a code (having a {\em skos:notation} property) and a category (having a {\em skos:prefLabel} property), 
then the property {\em disco:isValid} has to be stated indicating if the code is valid ('true') or missing ('false').

\subsection{Recommended Properties}

Which properties are not required but recommended within a particular context\footnote{{\em R-72-RECOMMENDED-PROPERTIES}}.
\textbf{{\em DISCO-C-XX:}}
The property ({\em skos:notation} is not mandatory for {\em disco:Variable}s, but recommended to indicate variable names.

\subsection{Value is Valid for Datatype}

Make sure that a value is valid for its datatype.
It has to be ensured, e.g., that a date is really a date, or that a xsd:nonNegativeInteger value is not negative. 
\textbf{{\em DISCO-C-XX:}}
Check if literal values of the property {\em disco:startDate} are really of the datatype {\em xsd:date}(not expressible in DL).

\subsection{Equivalent Properties}

\subsection{Functional Properties}

\subsection{Data Property Facets}

xsd:length of disco:purpose
